---
title: "Hydrochemical analysis of mid-low sections from North Patagonia Rivers, Argentina (2014-2015)"
author: "Vélez-Agudelo Camilo"
date: "18/2/2021"
output:
 html_document:
     toc: TRUE
     toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(gridExtra)
library(rstatix)
```

# **Exploratory Data Analysis** #

## Complete Data ##

These data base include measurements of summer and winter, as well as the coastal and continental sites

```{r }
varam <- readRDS("data.rds") ##original data##

varam.summer <- varam %>% 
    filter(Season == "summer") ##summer data##

varam.winter <- varam %>% 
    filter(Season == "winter")##winter data##
    
```


```{r echo = FALSE}
varam %>% 
    kable(caption = "Hidrochemical data (summer and winter) from Northen Patagonia River") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  font_size = 11,
                  position = "center",
                  full_width = T) %>%
    scroll_box(height = "300px", width = "900px")
```


```{r echo = FALSE}
summary.sta <- pastecs::stat.desc(
    varam[,c(4:19)], 
    )

summary.sta <- round(
    summary.sta, 2
    )

summary.sta %>% 
    kable(caption = "Statistical summary of physical and chemical variables (complete data)",
          align = "cccccccccccccccc") %>% 
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  font_size = 11,
                  position = "center",
                  full_width = T) %>%
    scroll_box(height = "300px", width = "900px")

```

### Bean Plot - Complete Data ###

Beanplots of physical and hydrochemical variables measured in the Colorado, Negro and Chubut Rivers. The dotted line is the overall mean for each variables while the solid line indicates the mean value in each season.

```{r echo = FALSE, fig.width = 10, fig.height = 6}

library(beanplot)

par(
    mar=c(2,2,2,2), 
    mfcol=c(3,5)
    )

beanplot(
    Temperature ~ Season*River,
    data = varam,
    bw = "nrd0",
    ll = 0.1,
    col = list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Temperature",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Conductivity ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Conductivity",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    pH ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="pH",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Salinity ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Salinity",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    HCO3 ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Bicarbonates",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Cl ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Chloride",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Hardness ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Hardness",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Ca ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Calcium",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Mg ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Magnesium",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    SO4 ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Sulphate",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Na ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Sodium",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    K ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Potassium",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )
legend(
    "right",
    fill=c("deepskyblue3","darkorange2"),
    legend=c("Winter","Summer"),
    bty="n",
    ncol=1
    )

beanplot(
    SiO2 ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Silica",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    NO3 ~ Season*River,
    data=varam,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Nitrate",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

```

### Bean Plot - In Inland Areas ###

The estuarine sites were removed from these graphics.

```{r echo = TRUE}

varam.inland <- varam[-c(1:6, 31:36, 67:75),]

```

```{r echo = FALSE, fig.width = 10, fig.height = 6}

par(
    mar=c(2,2,2,2), 
    mfcol=c(3,5)
    )

beanplot(
    Temperature ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Temperature",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Conductivity ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Conductivity",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    pH ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="pH",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Salinity ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Salinity",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    HCO3 ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Bicarbonates",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Cl ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Chloride",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Hardness ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Hardness",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Ca ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Calcium",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Mg ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Magnesium",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    SO4 ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Sulphate",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    Na ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Sodium",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    K ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Potassium",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )
legend(
    "right",
    fill=c("deepskyblue3","darkorange2"),
    legend=c("Winter","Summer"),
    bty="n",
    ncol=1
    )

beanplot(
    SiO2 ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Silica",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )

beanplot(
    NO3 ~ Season*River,
    data=varam.inland,
    bw="nrd0",
    ll=0.1,
    col=list("deepskyblue3",c("darkorange2","black")),
    horizontal=F,
    main="Nitrate",
    log="",
    side="both",
    beanlines="mean",
    show.names=T,
    innerborder=NA
    )
```


### Correlation Analysis - Complete Data ###




### Summer Data ##

These data include only the summer measurements doing in the three rivers.

```{r echo = FALSE}

summary.summer <- pastecs::stat.desc(
    varam.summer[,c(4:19)], 
    )

summary.summer <- round(
    summary.summer, 2
    )

summary.summer %>% 
    kable(caption = "Statistical summary of physical and chemical variables measured in summer",
          align = "cccccccccccccccc") %>% 
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  font_size = 11,
                  position = "center",
                  full_width = T) %>%
    scroll_box(height = "300px", width = "900px")
```


